---
title: "Fundamentals"
output: 
  html_document:
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Begin by installing and adding the following packages to your library:

```{r fundamentals-packs, echo = TRUE, eval = FALSE}
  install.packages(c("shiny", 
                     "shinythemes", 
                     "markdown", 
                     "leaflet", 
                     "tidyverse")) 
  library(shiny)
  library(shinythemes)
  library(markdown)
  library(leaflet)
  library(tidyverse)
```

Shiny apps are built on two main components:

  * the user interface (`ui.R`)
  * the server (`server.R`)
    
It is also possible to write apps in a single file called `app.R` that contains a user interface object and a server function. Whichever option you choose, note that the files **must** have these names. Today, we will write apps in single `app.R` files for simplicity. `app.R` must have the following form, noting the call to `shinyApp()` at the end:

```{r fundamentals-struct, echo = TRUE, eval = FALSE}
# define user interface
ui <- ...

# define server as function of input and output
server <- function(input, output) {...}
  
# call app
shinyApp(ui = ui, server = server)
```

Great! Now, let's build a simple app to get a feel for things. We will build an app that creates plots of time-series data from some of the built-in datasets in R. The code to produce the app is below, and a running version of it to the right. You can use the dropdown menu to select the dataset, and the radio buttons to select the color. 

## User interface

Behind the scenes of `ui` R is constructing HTML code that makes the front end of your app for your user. 

```{r fundamentals-ui, echo = TRUE, eval = FALSE}
ui <- pageWithSidebar(
  headerPanel('Plot selected time-series'),
  sidebarPanel(
    selectInput('dat', 
                'Dataset:',
                c("AirPassengers", "uspop", "mdeaths", "lynx", "USAccDeaths")
                ),
    radioButtons("color", "Plot color:",
                 choices = list("Red" = "red", 
                                "Orange " = "orange", 
                                "Yellow " = "yellow", 
                                "Green " = "green", 
                                "Blue " = "blue", 
                                "Indigo " = "darkblue", 
                                "Violet " = "violet"
                                ), 
                 selected = "blue"
                 )
  ),
  mainPanel(
    plotOutput('plot1'),
    textOutput("txtout")
  )
)
```

## Server

`server` talks to `ui` and (re)builds the R objects displayed in the `ui` based on user input.

Note the use of the function selectData outside the server function here. Why might we do this? We will return to this example when we discuss reactions.

```{r fundamentals-server, echo = TRUE, eval = FALSE}
server <- function(input, output) {
  # plot
  output$plot1 <- renderPlot({
    selected_data <- selectData(input$dat)
    plot(selected_data$series,
         col = input$color, ylab = input$dat, lwd = 2)
  })
  # description
  output$txtout <- renderText({
    selected_data <- selectData(input$dat)
    selected_data$description})
}

# outside function
selectData <- function(sent_var) {
  #nonchar <- input
  #iris[, c(input$xcol, input$ycol)]
  if (sent_var == "AirPassengers") {
    series <- datasets::AirPassengers
    description <- "The classic Box & Jenkins airline data. 
    Monthly totals of international airline passengers, 1949 to 1960."
    return(list(series = series, description = description))
  } else if (sent_var == "uspop") {
    series <- datasets::uspop
    description <- "This data set gives the population of the United States (in millions) 
    as recorded by the decennial census for the period 1790–1970."
    return(list(series = series, description = description))
  } else if (sent_var == "mdeaths") {
    series <- datasets::mdeaths
    description <- "Time series giving the monthly deaths from bronchitis, 
    emphysema and asthma in the UK, 1974–1979, males only."
    return(list(series = series, description = description))
  } else if (sent_var == "lynx") {
    series <- datasets::lynx
    description <- "Annual numbers of lynx trappings for 1821–1934 in Canada. 
    Taken from Brockwell & Davis (1991)"
    return(list(series = series, description = description))
  } else if (sent_var == "USAccDeaths") {
    series <- datasets::USAccDeaths
    description <- "A time series giving monthly totals of accidental deaths in the USA, 1973 - 1978."
    return(list(series = series, description = description))
  }
}
```

## Running your app

Make sure you have already run `library(shiny)` and then call:

```{r fundamentals-running, echo = TRUE, eval = FALSE}
# compile ui and server into an app
shinyApp(ui = ui, server = server)
# run the app!
runApp("our_first_app")
```

You can also do this in RStudio by hitting the big green arrow. `"our_first_app"` should be the name of your working directory.


