---
title: "Mapping"
output: 
  html_document:
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Like interactive plots, interactive maps are an attractive feature of Shiny and friends. A nice package for this is `Leaflet` (see [this](https://rstudio.github.io/leaflet/shiny.html)). We could spend a whole session on each of this package alone, but here is a very brief introduction to `Leaflet`. 

### Leaflet

Here is an example which renders a map of central Europe with randomly distributed points. Instead of `renderPlot()` use `renderLeaflet()` and associated arguments. 

```{r, eval = FALSE}
r_colors <- rgb(t(col2rgb(colors()) / 255))
names(r_colors) <- colors()

ui <- fluidPage(
  leafletOutput("mymap"),
  p(),
  actionButton("recalc", "New points")
)

server <- function(input, output) {
  points <- eventReactive(input$recalc, {
    cbind(rnorm(40) * 2 + 13, rnorm(40) + 48)
  }, ignoreNULL = FALSE)

  output$mymap <- renderLeaflet({
    leaflet() %>%
      addProviderTiles(providers$Stamen.TonerLite,
        options = providerTileOptions(noWrap = TRUE)
      ) %>%
      addMarkers(data = points())
  })
}

shinyApp(ui, server)
```
