---
title: "Widgets"
output: 
  html_document:
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ususally when you are building an RShiny app, it's because you want to allow users to provide inputs. Widgets are how you do that. 

There are three steps to creating output. First is telling `ui` what widget to create. If we want to create some checkboxes, that looks like:

```{r widgets-ui-input, echo = TRUE, eval = FALSE}
ui <- fluidPage(
  # make a group of checkboxes
  checkboxGroupInput("checkGroup", label = h3("Checkbox group"), 
    choices = list("Choice 1" = 1, "Choice 2" = 2, "Choice 3" = 3),
    selected = 1),
)
```

That's great, but the checkboxes won't do anything. Next we need to tell `server` what to render. The modification to `server` might look like: 

```{r widgets-serv, echo = TRUE, eval = FALSE}
server <- function(input, output) {
  output$value <- renderPrint({ input$checkGroup })
}
```

Note that internally, you reference the widget using `input$label_name`. This is different than the title, `input$title_name`, which is printed above the widget. Watch out for this, there are lots of names to keep track of when writing Shiny code. 

Finally, we go back to `ui` and tell it what to print. The complete `ui` might look something like:

```{r widgets-ui-output, echo = TRUE, eval = FALSE}
ui <- fluidPage(
  # make a group of checkboxes
  checkboxGroupInput("checkGroup", label = h3("Checkbox group"), 
    choices = list("Choice 1" = 1, "Choice 2" = 2, "Choice 3" = 3),
    selected = 1),
  # print a vertical space
  hr(),
  # print which boxes are selected
  fluidRow(column(3, verbatimTextOutput("value")))
)
```

That's the basic idea, but there are lots of different types of inputs and outputs. Let's look at some examples in the next tab. Much of this code is from RStudio's [Widget Gallery](https://shiny.rstudio.com/gallery/widget-gallery.html) which is an excellent resource.

### Types of Widgets

  * `actionButton`: action button (e.g "Submit" or "Run")
  * `checkboxGroupInput`:	a group of check boxes
  * `checkboxInput`:	a single check box
  * `dateInput`:	a calendar 
  * `dateRangeInput`:	a pair of calendars for selecting a date range
  * `fileInput`:	a file upload control wizard
  * `helpText`:	help text that can be added to an input form
  * `numericInput`: field to enter numbers
  * `radioButtons`: set of radio buttons
  * `selectInput`: box with choices to select from
  * `sliderInput`: slider bar
  * `submitButton`:	submit button
  * `textInput`: field to enter text

### Ways to render

  * `renderDataTable`:	DataTable
  * `renderImage`:	images (saved as a link to file)
  * `renderPlot`:	plots
  * `renderPrint`:	any printed output
  * `renderTable`:	data frame, matrix, other table like structures
  * `renderText`:	character strings
  * `renderUI`:	a Shiny tag object or HTML

### Ways to print

  * `dataTableOutput`:	DataTable
  * `htmlOutput`:	raw HTML
  * `imageOutput`:	image
  * `plotOutput`:	plot
  * `tableOutput`:	table
  * `textOutput`:	text
  * `uiOutput`:	raw HTML
  * `verbatimTextOutput`:	text, but ugly

Note that you should render and print the same object type. For example, you cannot `renderPlot` a `textOutput`.

